Word Learning as Network Growth: A Cross-linguistic Analysis

### 2018-11-12 journal/ms.tex
final journal draft

  \abstract{Children tend to produce words earlier when they are connected to a
variety of other words along both the phonological and semantic
dimensions. Though this connectivity effect has been extensively
documented, little is known about the underlying developmental
mechanism. One view suggests that learning is primarily driven by a
network growth model where highly connected words in the child's early
lexicon attract similar words. Another view suggests that learning is
driven by highly connected words in the external learning environment,
instead of highly connected words in the early internal lexicon. The
present study tests both scenarios systematically in both the
phonological and semantic domains across 10 languages. We show that
external connectivity in the learning environment drives growth in both
production- and comprehension-based vocabularies, even controlling for
word frequency and length. This pattern of findings suggests a word
learning mechanism where children harness their statistical learning
abilities to (indirectly) detect and learn highly connected words in the
learning environment.}
  \keywords{Language understanding; audio-visual processing; word learning; speech
perception; computational modeling. \\

@@

What factors shape vocabulary learning over the course of early
childhood? To investigate this question, scientists have adopted
multiple research strategies, from conducting controlled laboratory
experiments (e.g. Markman, 1990) to analyzing dense corpora capturing
language learning in context (e.g., B. C. Roy, Frank, DeCamp, Miller, \&
Roy, 2015). One strategy consists in documenting the timeline of words'
acquisition, and studying the properties that make words easy or hard to
learn. For example, within a lexical category, words that are more
frequent in child-directed speech are acquired earlier (J. C. Goodman,
Dale, \& Li, 2008). Other factors include word length, the mean length
of utterances in which the word occurs, and concreteness (see Braginsky,
Yurovsky, Marchman, \& Frank, 2016).

Besides these word-level properties, the structure of the lexicon (that
is, how words relate to each other) also predicts the age of acquisition
of words. The lexical structure can be characterized in terms of a
network where each node represents a word in the vocabulary, and each
link between two nodes represents a relationship between the
corresponding pair of words (e.g., Collins \& Loftus, 1975). Previous
studies have investigated early vocabulary structure by constructing
networks using a variety of word-word relations including shared
semantic features, target-cue relationships in free association norms,
co-occurrence in child directed speech, and phonological relatedness.
These studies have found that children tend to produce words that have
higher neighborhood density (i.e., high connectivity in the network)
earlier, both at the phonological and the semantic level (Carlson,
Sonderegger, \& Bane, 2014; Hills, Maouene, Riordan, \& Smith, 2010;
Hills, Maouene, Maouene, Sheya, \& Smith, 2009; Stella, Beckage, \&
Brede, 2017; Storkel, 2009).

@@

These two proposals represent two divergent ideas about the role of
lexical networks in acquisition. On the INT proposal, learning is driven
by known words with high connectivity to other known words (Figure 1,
left). Thus, the network structure is a causal factor in early word
learning, that is, children might be relying on the organization of
their past knowledge to determine future learning. For example, having a
rich and organized knowledge about the domain of dinosaurs would
facilitate the acquisition of new dinosaur-related words (Chi \& Koeske,
1983). In contrast, on the EXT approach, learning is driven by the
connectivity of words that are not known yet (Figure 1, right). Thus,
the relevant network structure is not internally represented by
children, and the related connectivity effect might be an epiphenomenon
of some low level properties of the linguistic input. For example,
highly connected words in the input could be more easily learned because
of their contextual diversity (e.g., Smith \& Yu, 2008) or because
caregivers emphasize such words in child-directed speech (MacWhinney,
2014).

@@

Hills et al. (2009) investigated the growth of lexico-semantic networks
in toddlers and found that growth did not proceed according to INT as
was originally hypothesised by Steyvers and Tenenbaum (2005), but rather
according to
EXT.\footnote{Besides INT and EXT, the authors tested a third mechanism (called the lure of associates) which resembles EXT in that it is driven by the connectivity of external nodes, except that this connectivity is computed with respect to words that are known. However, EXT is the externally-driven scenario that best predicted the data in this previous work.}
This is a profound finding because it suggests that learning in the
early stages is mostly driven by properties of the external input,
regardless of how past knowledge is organized. However, this work has
been done in the special case of networks that are based on 1) semantic
associations, 2) production-based vocabularies, and 2) data from
English-learning children, only. The extent to which this result depends
on the domain (e.g., semantic vs.~phonological connectivity), the
vocabulary measure (production vs.~comprehension) and culture/language
is still unclear. In this work, we test the generality of the finding
along these three dimensions.

First, we study the phonological network in addition to the semantic
network. These two networks represent different ways the mental lexicon
is structured. In particular, words that are neighbors in the semantic
network (e.g., \enquote{cat}, \enquote{dog}) are not necessarily
neighbors in the phonological network, and vice versa. Does the
phonological network also predict word learning? Previous work did find
an effect of words' connectivity in the phonological network on their
age of learning (Carlson et al., 2014; Stella et al., 2017; Storkel,
2009). In other words, words learned earlier in life tend to sound
similar to many other words than a word learned later in life. However,
this finding is \emph{a priori} compatible with both INT and EXT, and
previous studies did not explicitly compare these two mechanisms. Here,
we investigate whether phonological networks, like semantic networks,
grow through EXT, or if they rather grow via INT
(Figure~\ref{fig:growth}).

@@

Finally, we use developmental data in 10 languages. Lexical networks can
show more or less cross-linguistic variability along both the semantic
and phonological domains (Arbesman, Strogatz, \& Vitevitch, 2010; Youn
et al., 2016). Besides, cultures might differ in the way caregivers talk
to children (Cristia, Dupoux, Gurven, \& Stieglitz, 2017; Kuhl et al.,
1997), and this difference in the input could influence the children's
learning strategy. Thus, Cross-linguistic comparison is crucial to test
what mechanism is cognitively universal and is used by all children, and
what mechanism is specific to some patterns of learning that emerge due
to the particulars of a given language or culture (Bates \& MacWhinney,
1987; Slobin, 2014).

We test the growth scenarios using parent reports of children
vocabularies and their normative age of acquisition (Fenson et al.,
1994). Children may vary in their individual learning trajectories, but
the aggregate data leads to an average learning pattern which is highly
consistent. It is this normative trajectory that we model in the current
study, following the steps of previous research using similar datasets
(Braginsky et al., 2016; J. C. Goodman et al., 2008; Hills et al., 2010,
2009; Stella et al., 2017; Storkel, 2009)

The paper is organized as follows. First, we describe the datasets we
used and explain how we constructed the networks. Second, we analyze
static properties of words' connectivity in these networks (correlation
with AoA and shape of the distribution) and we explain how these
properties infom hypotheses about network growth. Next, we explicitly
fit the two hypothesized growth mechanisms to the data. We investigate
the extent to which the results obtained in Hills et al. (2009)
generalize to phonological networks and comprehension-based
vocabularies, and whether this generalization holds
cross-linguistically.

@@

We used data from Wordbank (Frank, Braginsky, Yurovsky, \& Marchman,
2017), an open repository aggregating cross-linguistic language
developmental data of the MacArthur-Bates Communicative Development
Inventory (CDI), a parent report vocabulary checklist. Parent report is
a reliable and valid measure of children's vocabulary that allows for
the cost-effective collection of datasets large enough to test
network-based models of acquisition (Fenson et al., 1994). When filling
out a CDI form, caregivers are either invited to indicate whether their
child \enquote{understands} (comprehension) or \enquote{understands and
says} (production) each of about 400-700 words. For younger children
(e.g., 8 to 18 months in the English data), both comprehension and
production are queried, whereas for older children (16 to 36 months)
only production is queried. Thus, we use data from younger children to
test comprehension and data from older children to test production
across 10 languages. Following previous studies (Hills et al., 2009;
Storkel, 2009), we restrict our analysis to nouns. Table \ref{tab:stats}
gives an overview of the data.

@@

\caption{\label{tab:stats}Statistics for dataset we used.}
\centering
\begin{tabular}[t]{>{\bfseries}lrlrl}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Comprehension} & \multicolumn{2}{c}{Production} \\
\cmidrule(l{2pt}r{2pt}){2-3} \cmidrule(l{2pt}r{2pt}){4-5}
Language & Nouns & Ages & Nouns & Ages\\
\midrule
Croatian & 209 & 8-16 & 312 & 16-30\\
Danish & 200 & 8-20 & 316 & 16-36\\
English & 209 & 8-18 & 312 & 16-30\\
French & 197 & 8-16 & 307 & 16-30\\
Italian & 209 & 7-24 & 312 & 18-36\\
Norwegian & 193 & 8-20 & 316 & 16-36\\
Russian & 207 & 8-18 & 314 & 18-36\\
Spanish & 208 & 8-18 & 312 & 16-30\\
Swedish & 205 & 8-16 & 339 & 16-28\\
Turkish & 180 & 8-16 & 297 & 16-36\\
\bottomrule
\end{tabular}
\end{table}

@@

We constructed semantic networks for English data following the
procedure outlined in Hills et al. (2009), as follows. We used as an
index of semantic relatedness the Florida Free Association Norms
(Nelson, McEvoy, \& Schreiber, 1998). This dataset was collected by
giving adult participants a word (the cue), and asking them to write the
first word that comes to mind (the target). For example, when given the
word \enquote{ball}, they might answer with the word \enquote{game}. A
pair of nodes were connected by a directed link from the cue to the
target if there was a cue-target relationship between these nodes in the
association norms. The connectivity of a given node was characterized by
its \emph{indegree}: the number of links for which the word was the
target. To model growth from month to month, we constructed a different
network at each month, based on the words that have been acquired by
that month.

Since the free association norms are available only in English, we used
the hand-checked translation equivalents available in Wordbank, which
allowed us to use the English association norms across languages. Using
the same association data across languages does not necessarily mean
that the resulting networks will be the same across languages, or that
these networks will grow similarly. Indeed, though this approximation
assumes that the semantic similarity measure is universal----which is a
reasonable assumption (e.g., Youn et al., 2016), the set of words
acquired by children as well as the timeline of this acquisition can
still vary from language to language leading to possibility different
learning strategies.

@@

In previous studies, two nodes were linked if they had an edit distance
of 1 (Carlson et al., 2014; Stella et al., 2017; Storkel, 2009).
However, these studies reported a contribution of phonological networks
to word learning when these networks were built using a rich adult
vocabulary. However, since the focus of the current study is on the
mechanism of growth, the networks should be based on the children's
early vocabulary which, nevertheless, contains very few word pairs with
an edit distance of 1. Thus, we increased the threshold from 1 to 2,
that is, two nodes were related if their edit distance was equal to 1 or
2.\footnote{We also considered the case of an edit distance of 1 as well as the continuous measure, i.e., the inverse edit distance without threshold. In both cases, the results were weaker that those done with a threshold of 2.}
The connectivity of a given node was characterized with its
\emph{degree}: the number of links it shares with other words.

@@

We start by analyzing word connectivity in the global (static) network.
We constructed this network using nouns learned by the oldest age for
which we have CDI data (e.g., in English this corresponds, in
comprehension, to the network by 18 months, and in production, to the
network by 30 months). This global network is the end-state towards
which both INT and EXT converge by the last month of learning. Moreover,
following Hills et al. (2009), we used this end-state network as a proxy
for the external connectivity in the learning environment. Below we
analyze properties of this global networks that are relevant to INT
and/or EXT. In order to compare various predictors on the same data, we
restrict the analysis to the subset of nouns for which we had both
semantic and phonological information in each language.

@@

Connectivity in the global network is directly related to EXT as it
represents the explicit criterion this growth scenario uses to determine
what words should be learned first (Figure \ref{fig:growth}). Therefore,
a direct consequence of an EXT-like growth scenario is a correlation
between connectivity in the global network and the age of
acquisition.\footnote{This correlation is also compatible with INT, although the causality is reversed. Indeed, from the perspective of this growth scenario, higher connectivity in the global network is caused by earlier learning, not the other way around. Some words end up being highly connected in the global network precisely because they happen to be acquired earlier and, therefore, have a higher chance of accumulating more links over time.}
Figure \ref{fig:corrAll} shows how the age of production and
comprehension for each word varies as a function of its degree (or
indegree for the semantic networks) as well as the correlation values.
For ease of visual comparison, the predictor (i.e., the degree) was
centered and scaled across languages.

The plots show, overall, a negative correlation between the month of
acquisition and the degree. In production data, the average correlation
across languages was -0.24 (SD=0.10) for the semantic networks and -0.30
(SD=0.05) for the phonological networks. In comprehension data, the
average correlation was -0.21 (SD=0.08) for the semantic networks and
-0.21 (SD=0.07) for the phonological networks. These results indicate
that nouns with higher degrees are generally learned earlier, thus
replicating previous findings in English (e.g., Storkel 2004, 2009;
Hills et al. 2009) and extending these findings to ten different
languages, generally, in both production- and comprehension-based
vocabularies.

@@

We also analyzed the global network's degree distribution. The shape of
this distribution is particularly relevant to INT as this growth
scenario is known to generate networks with a power-law degree
distribution, i.e., a distribution of the form
\(p(k) \propto \frac{1}{k^{\alpha}}\) (Barabasi \& Albert, 1999). If the
network displays this property, this fact would suggest, but not prove,
an INT-like generative process. If the degree distribution does not
follow a power law, this would largely weaken the case for INT. The
log-log plots are shown in Figure \ref{fig:degreeDist}. We fit a power
law to each empirical degree distribution following the procedure
outlined in Clauset, Shalizi, and Newman (2009) and using a related R
package (poweRlaw, Gillespie, 2015).

@@

Overall, we could not reject the null hypothesis of a power-law
distribution: The \(p\)-value was generally above 0.1 in almost all
languages for both production and comprehension. That said, phonological
networks had relatively larger cut-offs than semantic networks. As was
suggested by Arbesman et al. (2010), these \enquote{truncated}
power-laws in phonological networks--as well as the observed
cross-linguistic variability in the value of the cut-offs--may reflect
the various constraints that exist on word formation such as the number
of phonemes in the language, the phonotactics (i.e., the way sound
sequences are arranged in words), and the length of words. Such
constraints may limit the number of words that are phonologically
similar, thus leading to distributions which decay faster than a
non-truncated power law.

In sum, the static properties of the global network are \emph{a priori}
compatible with both INT and EXT. In order to decide between these two
developmental scenarios, we need to fit explicit growth models to the
data.

@@

As for the factors of interest, i.e., semantic and phonological
connectivity, we found cross-linguistic differences. Connectivity
contributes to learning in some languages but not in other. In
particular, semantic connectivity does not explain variance in English
data beyond that explained by phonological connectivity, frequency and
length. This contrasts with the original finding in Hills et al. (2009).
However, this might be due to our using of a slightly different model
(which included word length as a covariate) and a larger dataset. That
said, and despite these apparent cross-linguistic differences, both
phonological and semantic connectivity are significant predictors in the
combined model.

@@

One possibility is that these patterns emerge from children's use of
statistical learning abilities (Aslin \& Newport, 2012; Saffran, Aslin,
\& Newport, 1996; Smith \& Yu, 2008). The term \enquote{statistical
learning} has been used in the developmental literature to describes the
process by which one acquires information about their environment
through keeping track of the frequency distribution of some elements
(e.g., words) in different contexts. An important property of this kind
of learning is that it occurs without explicit instructions and through
mere exposure to the input. Previous work in the line of research has
documented mechanisms which can explain the patterns found in the
current study.

For example, in the semantic domain, growth according to EXT can be
explained by a mechanism similar to cross-situational learning (Pinker,
2013; Smith \& Yu, 2008; Yurovsky \& Frank, 2015). According to this
mechanism, children track the co-occurrence of concrete nouns with their
possible semantic referents. The referent of a word heard in only one
naming situation can be ambiguous (e.g., when the word \enquote{ball} is
heard for the first time in the presence of both a ball and a chair),
but hearing the same word in a diversity of semantic contexts allows the
learner to narrow down the set of possible word-object mappings. In our
case, free association is related to contextual co-occurrence
(Griffiths, Steyvers, \& Tenenbaum, 2007), meaning that highly connected
words will tend to occur in a variety of speech and referential
contexts. This fact makes such words easier to learn because they have
more referential disambiguating cues across learning contexts, and
crucially, even without knowing the entire set of words with witch they
co-occur (hence the similarity with EXT). This possibility is supported
by the finding that words' diversity of occurrence in child directed
speech predicts their age of learning (Hills et al., 2010).

In the phonological case, network growth according to EXT is also
compatible with a scenario whereby children are tracking lwo level
statistical patterns, e.g., high probability sound sequences. Indeed,
connectivity in the phonological network is inherently correlated with
phonotactic probability (M. S. Vitevitch, Luce, Pisoni, \& Auer, 1999).
That is, highly connected words tend to be made of frequent sound
sequences. Children are sensitive to local phonotactic regularities
(Jusczyk, Luce, \& Charles-Luce, 1994) and this sensitivity might lead
them to learn higher-probability words more easily (Storkel, 2001). This
explanation is supported by computational simulations that shows how
learning general phonotactics patterns create ``well-worn paths'' which
allow the models to represent several distinct but phonologically
neighboring words (Dell, Juliano, \& Govindjee, 1993; Takac, Knott, \&
Stokes, 2017).

@@

This work shares a number of limitations with previous studies using
similar research strategy and datasets. Chief among these limitations is
the fact that the age of word acquisition is computed using different
children at different ages (due to the fact that available CDI data is
mainly cross-sectional). Although this measure has proven highly
consistent (Fenson et al., 1994), it led us to focus on studying the
learning mechanism of the \enquote{average} child. Individual
trajectories, however, could show different learning patterns. For
example, using longitudinal data Beckage, Smith, and Hills (2011) found
differences between typical and late talkers in terms of the semantic
network structure. Besides, although our study endorses the
externally-driven account of network growth, this does not mean
individual children never use some variant of INT or some combination of
both INT and EXT (Beckage and Colunga, under review). For example, some
children develop \enquote{islands of expertise}, that is, well organized
knowledge about a certain topic (e.g., birds or dinosaurs). This prior
knowledge enables these children to learn new related words more easily
(e.g., Chi \& Koeske, 1983).

>>>>

  \abstract{Children tend to produce words earlier when they are connected to a
variety of other words along both the phonological and semantic
dimensions. Though this connectivity effect has been extensively
documented, little is known about the underlying developmental
mechanism. One view suggests that learning is primarily driven by a
network growth model where highly connected words in the child's early
lexicon attract similar words. Another view suggests that learning is
driven by highly connected words in the external learning environment,
instead of highly connected words in the early internal lexicon. The
present study tests both scenarios systematically in both the
phonological and semantic domains across 10 languages. We show that
external connectivity in the learning environment drives growth in both
production- and comprehension-based vocabularies, even controlling for
word frequency and length. This pattern of findings suggests a word
learning mechanism where children harness their statistical learning
abilities to detect and learn highly connected words in the learning
environment.}
  \keywords{Language understanding; audio-visual processing; word learning; speech
perception; computational modeling. \\

@@

What factors shape vocabulary learning over the course of early
childhood? To investigate this question, scientists have adopted
multiple research strategies, from conducting controlled laboratory
experiments (e.g. Markman, 1990) to analyzing dense corpora capturing
language learning in context (e.g., B. C. Roy, Frank, DeCamp, Miller, \&
Roy, 2015). One strategy consists in documenting the timeline of words'
acquisition, and studying the properties that make words easy or hard to
learn. For example, within a lexical category, words that are more
frequent in child-directed speech are acquired earlier (J. C. Goodman,
Dale, \& Li, 2008). Other factors include word length and the mean
length of utterances in which the word occurs (e.g., Braginsky,
Yurovsky, Marchman, \& Frank, 2016; Swingley \& Humphrey, 2018).

Besides these word-level properties, the structure of the lexicon (that
is, how words relate to one another) also predicts the Age of
Acquisition (AoA) of words. The lexical structure can be characterized
in terms of a network where each node represents a word in the
vocabulary, and each link between two nodes represents a relationship
between the corresponding pair of words (e.g., Collins \& Loftus, 1975).
Previous studies have investigated early vocabulary structure by
constructing networks using a variety of word-word relations including
shared semantic features (McRae, Cree, Seidenberg, \& McNorgan, 2005),
target-cue relationships in free association norms (Nelson, McEvoy, \&
Schreiber, 1998), co-occurrence in child directed speech (MacWhinney,
2014), and phonological relatedness (Vitevitch, 2008). These studies
have generally found that children tend to produce words that have
higher neighborhood density (i.e., high connectivity in the network)
earlier, both at the phonological and the semantic level (Carlson,
Sonderegger, \& Bane, 2014; Hills, Maouene, Riordan, \& Smith, 2010;
Hills, Maouene, Maouene, Sheya, \& Smith, 2009; Stella, Beckage, \&
Brede, 2017; Storkel, 2009).

@@

These two proposals represent two divergent ideas about the role of
lexical networks in acquisition. On the INT proposal, learning is driven
by known words with high connectivity to other known words (Figure 1,
left). Thus, the network structure is a causal factor in early word
learning, that is, children rely on the organization of their past
knowledge to determine future learning. For example, having a rich and
organized knowledge about the domain of dinosaurs would facilitate the
acquisition of new dinosaur-related words (Chi \& Koeske, 1983). In
contrast, on the EXT approach, learning is driven by the connectivity of
words that are not known yet (Figure 1, right). Thus, the relevant
network structure is not internally represented by children, and the
observed connectivity effect might be an epiphenomenon of some
properties of the linguistic input. For example, highly connected words
in the input could be more easily learned because of their contextual
diversity (e.g., Smith \& Yu, 2008) or because caregivers emphasize such
words in child-directed speech (MacWhinney, 2014).

@@

Hills et al. (2009) investigated the growth of lexico-semantic networks
in toddlers and found that growth did not proceed according to INT as
was originally hypothesised by Steyvers and Tenenbaum (2005), but rather
according to
EXT.\footnote{Besides INT and EXT, the authors tested a third mechanism (called the lure of associates) which resembles EXT in that it is driven by the connectivity of external nodes, except that this connectivity is computed with respect to words that are known. However, EXT is the externally-driven scenario that best predicted the data in this previous work.}
This is a profound finding because it suggests that learning in the
early stages is mostly driven by properties of the external input,
regardless of how past knowledge is organized. However, this work has
been done in the special case of networks that are based on 1) semantic
associations, 2) production-based vocabularies, and 3) data from
English-learning children, only. The extent to which this result depends
on the domain (e.g., semantic vs.~phonological connectivity), the
vocabulary measure (production vs.~comprehension) and culture/language
is still unclear. In this work, we test the generality of the finding
along these three dimensions.

First, we study the phonological network in addition to the semantic
network. These two networks represent different ways the mental lexicon
is structured (Beckage \& Colunga, 2016). In particular, words that are
neighbors in the semantic network (e.g., \enquote{cat}, \enquote{dog})
are not necessarily neighbors in the phonological network, and vice
versa. Does the phonological network also predict word learning?
Previous work did find an effect of words' connectivity in the
phonological network on their age of learning (Carlson et al., 2014;
Stella et al., 2017; Storkel, 2009). In other words, words learned
earlier in life tend to sound similar to many other words than a word
learned later in life. However, this finding is \emph{a priori}
compatible with both INT and EXT, and previous studies did not
explicitly compare these two mechanisms. Here, we investigate whether
phonological networks, like semantic networks, grow through EXT, or if
they rather grow via INT (Figure~\ref{fig:growth}).

@@

Finally, we use developmental data in 10 languages. Lexical networks can
show more or less cross-linguistic variability along both the semantic
and phonological domains (Arbesman, Strogatz, \& Vitevitch, 2010; Youn
et al., 2016). Besides, cultures might differ in the way caregivers talk
to children (Cristia, Dupoux, Gurven, \& Stieglitz, 2017; Kuhl et al.,
1997), and this difference in the input could influence the way in which
the children's networks grow. Thus, Cross-linguistic comparison is
crucial to test what growth mechanism is cognitively universal and is
used by all children, and what mechanism is specific to some patterns of
learning that emerge due to the particulars of a given language or
culture (Bates \& MacWhinney, 1987; Slobin, 2014).

We test the growth scenarios using parent reports of children
vocabularies and their normative age of acquisition (Fenson et al.,
1994). Children may vary in their individual learning trajectories, but
the aggregate data leads to an average learning pattern which is highly
consistent. It is this normative trajectory that we model in the current
study, following in this regard the steps of previous research using
similar datasets (Braginsky et al., 2016; J. C. Goodman et al., 2008;
Hills et al., 2010, 2009; Stella et al., 2017; Storkel, 2009)

The paper is organized as follows. First, we describe the datasets we
used and explain how we constructed the networks. Second, we analyze
static properties of words' connectivity in these networks (correlation
with age of acquisition and shape of the distribution) and we explain
how these properties infom hypotheses about network growth. Next, we
explicitly fit the two hypothesized growth mechanisms to the data. We
investigate the extent to which the results obtained in Hills et al.
(2009) generalize to phonological networks and comprehension-based
vocabularies, and whether this generalization holds
cross-linguistically.

@@

We used data from Wordbank (Frank, Braginsky, Yurovsky, \& Marchman,
2017), an open repository aggregating cross-linguistic language
developmental data of the MacArthur-Bates Communicative Development
Inventory (CDI), a parent report vocabulary checklist. Parent report is
a reliable and valid measure of children's vocabulary that allows for
the cost-effective collection of datasets large enough to test
network-based models of acquisition (Fenson et al., 1994). When filling
out a CDI form, caregivers are either invited to indicate whether their
child \enquote{understands} (comprehension) or \enquote{understands and
says} (production) each of about 400-700 words. For younger children
(e.g., 8 to 18 months in the English data), both comprehension and
production are queried, whereas for older children (16 to 36 months)
only production is queried. Due to these limitations, we use data from
younger children to test comprehension and data from older children to
test production. Following previous studies (Hills et al., 2009;
Storkel, 2009), we restrict our analysis to nouns. Table \ref{tab:stats}
gives an overview of the data we used.

@@

\caption{\label{tab:stats}Statistics for the dataset we used.}
\centering
\begin{tabular}[t]{>{\bfseries}lrlrl}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Comprehension} & \multicolumn{2}{c}{Production} \\
\cmidrule(l{2pt}r{2pt}){2-3} \cmidrule(l{2pt}r{2pt}){4-5}
Language & Nouns & Ages & Nouns & Ages\\
\midrule
Croatian & 209 & 8-16 & 312 & 16-30\\
Danish & 200 & 8-20 & 316 & 16-36\\
English & 209 & 8-18 & 312 & 16-30\\
French & 197 & 8-16 & 307 & 16-30\\
Italian & 209 & 7-24 & 312 & 18-36\\
Norwegian & 193 & 8-20 & 316 & 16-36\\
Russian & 207 & 8-18 & 314 & 18-36\\
Spanish & 208 & 8-18 & 312 & 16-30\\
Swedish & 205 & 8-16 & 339 & 16-28\\
Turkish & 180 & 8-16 & 297 & 16-36\\
\bottomrule
\end{tabular}
\end{table}

@@

We constructed semantic networks for English data following the
procedure outlined in Hills et al. (2009), as follows. We used as an
index of semantic relatedness the Florida Free Association Norms (Nelson
et al., 1998). This dataset was collected by giving adult participants a
word (the cue), and asking them to write the first word that comes to
mind (the target). For example, when given the word \enquote{ball}, they
might answer with the word \enquote{game}. A pair of nodes were
connected by a directed link from the cue to the target if there was a
cue-target relationship between these nodes in the association norms.
The connectivity of a given node was characterized by its
\emph{indegree}: the number of links for which the word was the target.
To model growth from month to month, we constructed a different network
at each month, based on the words that have been acquired by that month.

Since the free association norms are available only in English, we used
the hand-checked translation equivalents available in Wordbank, which
allowed us to use the English association norms across 10 languages.
Using the same association data across languages does not necessarily
lead to similar networks. Indeed, though this approximation assumes that
the semantic similarity measure is universal--which is a reasonable
approximation (e.g., Youn et al., 2016), the set of words acquired by
children as well as the timeline of this acquisition can still vary from
language to language leading to possibility different learning
strategies.

@@

In previous studies, two nodes were linked if they had an edit distance
of 1 (Carlson et al., 2014; Stella et al., 2017; Storkel, 2009).
However, these studies reported a contribution of phonological
connectivity to word learning when networks were built using a rich
adult vocabulary. Since the focus of the current study is on the
mechanism of growth, the networks should be based on the children's
early vocabulary. The latter, nevertheless, contains very few word pairs
with an edit distance of 1. Thus, we increased the threshold from 1 to
2, that is, two nodes were related if their edit distance was equal to 1
or
2.\footnote{We also considered the case of an edit distance of 1 as well as the continuous measure, i.e., the inverse edit distance without threshold. In both cases, the results were weaker that those obtained with a threshold of 2.}
The connectivity of a given node was characterized with its
\emph{degree}: the number of links it shares with other words.

@@

We start by analyzing word connectivity in the global (static) network.
We constructed this network using nouns learned by the oldest age for
which we have CDI data (e.g., in English this corresponds, in
comprehension, to the network by 18 months, and in production, to the
network by 30 months). This global network is the end-state towards
which both INT and EXT converge by the last month of learning. Moreover,
following Hills et al. (2009), we used this end-state network as a proxy
for the external connectivity in the learning environment. Below we
analyze properties of this global networks that may a priori hint at an
INT- or EXT-like growth. In order to compare various predictors on the
same data, we restrict the analysis to the subset of nouns for which we
had both semantic and phonological information in each language.

@@

Connectivity in the global network is directly related to EXT as it
represents the explicit criterion this growth scenario uses to determine
what words should be learned first (Figure \ref{fig:growth}). Therefore,
a direct consequence of an EXT-like growth scenario is a correlation
between connectivity in the global network and the age of acquisition.
This correlation is also compatible with INT, although the causality is
reversed: higher connectivity in the global network is caused by earlier
learning, not the other way around. Some words end up being highly
connected in the global network precisely because they happen to be
acquired earlier and, therefore, have a higher chance of accumulating
more links over time. Thus, the correlates between connectivity in the
end-state network and AoA is an important blusprint of both EXT and INT.
If there is no such correlation, neither growth scenarios can be posited
as a learning mechanism.

Figure \ref{fig:corrAll} shows how the age of production and
comprehension for each word varies as a function of its degree (or
indegree for the semantic networks) as well as the correlation values.
For ease of visual comparison, the predictor (i.e., the degree) was
centered and scaled across languages. The plots show, overall, a
negative correlation between the month of acquisition and the degree. In
production data, the average correlation across languages was -0.24
(SD=0.10) for the semantic networks and -0.30 (SD=0.05) for the
phonological networks. In comprehension data, the average correlation
was -0.21 (SD=0.08) for the semantic networks and -0.21 (SD=0.07) for
the phonological networks. These results indicate that nouns with higher
degrees are generally learned earlier, thus replicating previous
findings in English (e.g., Storkel 2004, 2009; Hills et al. 2009) and
extending these findings to 10 different languages, generally, in both
production- and comprehension-based vocabularies.

@@

We also analyzed the global network's degree distribution. The shape of
this distribution is particularly relevant to INT as this growth
scenario is known to generate networks with a power-law degree
distribution, i.e., a distribution of the form
\(p(k) \propto \frac{1}{k^{\alpha}}\) (Barabasi \& Albert, 1999). If the
end-state network displays this property, this fact would suggest, but
not prove, an INT-like generative process. If, however, the degree
distribution is very different from a power law, this would
significantly weaken the case for INT. The log-log plots are shown in
Figure \ref{fig:degreeDist}. We fit a power law to each empirical degree
distribution following the procedure outlined in Clauset, Shalizi, and
Newman (2009) and using a related R package (poweRlaw, Gillespie, 2015).

@@

Overall, we could not reject the null hypothesis of a power-law
distribution: The \(p\)-value was generally above 0.1 in almost all
languages for both production and comprehension. That said, phonological
networks had relatively larger cut-offs than semantic networks. As was
suggested by Arbesman et al. (2010), these \enquote{truncated}
power-laws in phonological networks --as well as the cross-linguistic
variability in the value of the cut-offs-- may reflect the various
constraints that exist on word formation such as the number of phonemes
in the language, the phonotactics (i.e., the way sound sequences are
arranged in words), and the length of words. Such constraints may limit
the number of words that are phonologically similar, thus leading to
distributions which decay faster than a non-truncated power law.

In sum, the static properties of the end-state network are \emph{a
priori} compatible with both INT and EXT. In order to decide between
these two developmental scenarios, we need to fit explicit growth models
to the data.

@@

As for the factors of interest, i.e., semantic and phonological
connectivity, we found cross-linguistic differences. Connectivity
contributes to learning in some languages but not in other. In
particular, semantic connectivity does not explain variance in English
data beyond that explained by phonological connectivity, frequency and
length. This contrasts with the original finding in Hills et al. (2009).
However, this might be due to our using a slightly different model
(which included word length as a covariate) and a larger dataset. That
said, and despite these apparent cross-linguistic differences, both
phonological and semantic connectivity are significant predictors in the
combined model (Figure \ref{fig:staticAll}).

@@

One possibility is that these patterns emerge from children's use of
statistical learning abilities (Aslin \& Newport, 2012; Saffran, Aslin,
\& Newport, 1996; Smith \& Yu, 2008). The term \enquote{statistical
learning} has been used in the developmental literature to describes the
process by which one acquires information about their environment
through tracking the frequency distribution of some elements (e.g.,
words) in different contexts. An important property of this kind of
learning is that it occurs without explicit instructions and through
mere exposure to the input. Previous work in the line of research has
documented specific mechanisms which can explain the patterns found in
the current study.

For example, in the semantic domain, growth according to EXT can be
explained by a mechanism similar to cross-situational learning (Pinker,
2013; Smith \& Yu, 2008; Yurovsky \& Frank, 2015). According to this
mechanism, children track the co-occurrence of concrete nouns with their
possible semantic referents. The referent of a word heard in only one
naming situation can be ambiguous (e.g., when the word \enquote{ball} is
heard for the first time in the presence of both a ball and a chair),
but hearing the same word in a diversity of semantic contexts allows the
learner to narrow down the set of possible word-object mappings. In our
case, free association (used to determine semamtnic network
connectivity) is related to contextual co-occurrence (Fourtassi \&
Dupoux, 2013; Griffiths, Steyvers, \& Tenenbaum, 2007), meaning that
highly connected words will tend to occur in a variety of speech and
referential contexts. This fact makes such words easier to learn because
they have more referential disambiguating cues across learning contexts.
Crucially, children can learn these words without knowing the other
words with witch they co-occur (hence the similarity with EXT). This
possibility is supported by the finding that words' diversity of
occurrence in child directed speech predicts their age of learning
(Hills et al., 2010).

In the phonological case, network growth according to EXT is also
compatible with a scenario whereby children are tracking lwo level
statistical patterns, e.g., high probability sound sequences. Indeed,
connectivity in the phonological network is inherently correlated with
phonotactic probability (Vitevitch, Luce, Pisoni, \& Auer, 1999). That
is, highly connected words tend to be made of frequent sound sequences.
Children are sensitive to local phonotactic regularities (Jusczyk, Luce,
\& Charles-Luce, 1994) and this sensitivity might lead them to learn
higher-probability words more easily (Storkel, 2001). This explanation
is supported by computational simulations that show how learning general
phonotactics patterns create ``well-worn paths'' which allow the models
to represent several distinct but phonologically neighboring words
(Dell, Juliano, \& Govindjee, 1993; Takac, Knott, \& Stokes, 2017).

@@

This work shares a number of limitations with previous studies using
similar research strategy and datasets. Chief among these limitations is
the fact that the age of word acquisition is computed using different
children at different ages (due to the fact that available CDI data is
mainly cross-sectional). Although this measure has proven highly
consistent (Fenson et al., 1994), it led us to focus on studying the
learning mechanism of the \enquote{average} child. Individual
trajectories, however, could lead to different netwtok strcutures and
show different learning patterns. For example, using longitudinal data
Beckage, Smith, and Hills (2011) found differences between typical and
late talkers in terms of the semantic network structure. Besides,
although our study endorses the externally-driven account of network
growth, this does not mean individual children never use some variant of
INT or some combination of both INT and EXT (Beckage and Colunga, under
review). For example, some children develop \enquote{islands of
expertise}, that is, well organized knowledge about a certain topic
(e.g., birds or dinosaurs). This prior knowledge enables these children
to learn new related words more easily (e.g., Chi \& Koeske, 1983).
